---
title: "Compare observed and model dune topography"
author: "Femke Batsleer"
date: "14 april 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load data

Field data, with predictions with INLA; interpolations with INLA  
data from IBM model

```{r message=FALSE, warning=FALSE}
#packages
library(tidyr)
library(dplyr)
library(INLA)
library(plotly)
library(forecast)
library(gridExtra)

#field data and predictions from INLA
lidar_data <- read.csv("output/Data_and_predictions_INLA_fielddata.csv", sep=",")
interlidar_data <- read.csv("output/interpolations_INLA_fielddata.csv", sep=",")

#data IBM output
ibm_data <- read.csv("data/data_model.csv", sep=",") %>% slice(-836) %>% #delete wrong (double) line
  filter(error < 0.4) %>% #filter outlier from error
  mutate(rep = as.factor(rep))
```

#Transform variables

To get comparable values/units and scale them the same  


```{r}
#summary(lidar_data)
#summary(ibm_data)

#divide JC in two, than comparable to field data
ibm_data <- ibm_data %>% mutate(JC2 = JC/2) %>% 
  #standardize IBM data: P is from 0 to 1, so ok; Hstd scaled JC2
  mutate(Pstd = P, Hstd = scale(JC2, center=mean(JC2), scale=sd(JC2)))
```

#Models for IBM data

```{r}
#first, get data for interpolations
interibm_data <- tibble(Pstd=seq(0,1, by=0.01), H=seq(20,70,by=0.5)) %>%
  tidyr::expand(Pstd, H) %>%
  mutate(Hstd=scale(H, center=mean(ibm_data$JC2), scale=sd(ibm_data$JC2))) %>% #scale by IBM data, to get similar values
  mutate(volume = NA, error = NA)#add empty values for volume and error
  
#select needed columns from ibm_data
ibm_all <- ibm_data %>% select(Pstd, Hstd, volume, error) %>%
  #bind with interpol data to feed it to INLA
  bind_rows(select(interibm_data, -H))
  
#INLA
#for volume
f0_diff <- volume ~  Pstd + Hstd + Pstd:Hstd +
  I(Pstd^2) + I(Hstd^2) +Hstd:I(Pstd^2) + Pstd:I(Hstd^2)
M0_diff <- inla(f0_diff, 
                #control.compute = list(dic = TRUE, waic = TRUE),
                control.predictor = list(compute = TRUE),
                family = "gaussian", data = ibm_all)
summary(M0_diff)

#for error
f0_err <- error ~ Pstd + Hstd + Pstd:Hstd +
  I(Pstd^2) + I(Hstd^2) +Hstd:I(Pstd^2) + Pstd:I(Hstd^2)
M0_err <- inla(f0_err,
               control.predictor = list(compute = TRUE),
               family = "gamma", data = ibm_all)
summary(M0_err)

#save(M0_diff, file="output/M0_diff.rda")
#save(M0_err, file="output/M0_err.rda")

#extract interpolation data from IBM models
M0_diff_inter <- M0_diff$summary.fitted.values %>%
  slice(990:n()) %>% #select predictions from dataset that are not input data
  select(mean, sd, `0.025quant`, `0.975quant`) %>% #select neaded values
  #rename to get similar values as in interlidar_data
  rename(Int_diff_mean = mean, Int_diff_sd = sd, Int_diff_0.025quant=`0.025quant`, Int_diff_0.975quant = `0.975quant`)

M0_err_inter <- M0_err$summary.fitted.values %>%
  slice(990:n()) %>% #select predictions from dataset that are not input data
  select(mean, sd, `0.025quant`, `0.975quant`) %>% #select neaded values
  #rename to get similar values as in interlidar_data
  rename(Int_err_mean = mean, Int_err_sd = sd, Int_err_0.025quant=`0.025quant`, Int_err_0.975quant = `0.975quant`)



#bind these interpolations/predictions to interibm_data
interibm_data <- interibm_data %>% bind_cols(M0_diff_inter) %>% bind_cols(M0_err_inter)
```

#Effect sizes

Visualisation of the effect sizes/signs of the covariates

```{r fig.height=10, fig.width=12}
#load INLA models from lidar data
load("output/M1_diff.rda")
load("output/M1_err.rda")

#lidar data get coefficients from models
post_beta1_diff <- M1_diff$summary.fixed[, c("mean", "0.025quant", "0.975quant")]
post_beta1_err <- M1_err$summary.fixed[, c("mean", "0.025quant", "0.975quant")]

post_beta1_diff <- post_beta1_diff %>% tibble::rownames_to_column(var="WhichVar") %>%
  mutate(WhichVar = factor(WhichVar, levels=c("meandVstd", "Diststd",  "Pstd:I(Hstd^2)", "Hstd:I(Pstd^2)",
                                              "I(Hstd^2)","I(Pstd^2)", "Pstd:Hstd", "Hstd", "Pstd","Intercept" ))) %>%
  mutate(WhichVar = recode(WhichVar, "meandVstd"="meandV", "Diststd"="Dist",
                           "Pstd:I(Hstd^2)"="P:JC²", "Hstd:I(Pstd^2)"="JC:P²",
                           "I(Hstd^2)"="JC²","I(Pstd^2)"="P²", "Pstd:Hstd"="P:JC", "Hstd"="JC",
                           "Pstd"="P","(Intercept)"="(Intercept)" ))

post_beta1_err <- post_beta1_err %>% tibble::rownames_to_column(var="WhichVar") %>%
  mutate(WhichVar = factor(WhichVar, levels=c("meandVstd", "Diststd",  "Pstd:I(Hstd^2)", "Hstd:I(Pstd^2)",
                                              "I(Hstd^2)","I(Pstd^2)", "Pstd:Hstd", "Hstd", "Pstd","Intercept" )))  %>%
  mutate(WhichVar = recode(WhichVar, "meandVstd"="meandV", "Diststd"="Dist",
                           "Pstd:I(Hstd^2)"="P:JC²", "Hstd:I(Pstd^2)"="JC:P²",
                           "I(Hstd^2)"="JC²","I(Pstd^2)"="P²", "Pstd:Hstd"="P:JC", "Hstd"="JC",
                           "Pstd"="P","(Intercept)"="(Intercept)" ))

#ibm data get coefficients from models
post_beta0_diff <- M0_diff$summary.fixed[, c("mean", "0.025quant", "0.975quant")]
post_beta0_err <- M0_err$summary.fixed[, c("mean", "0.025quant", "0.975quant")]

post_beta0_diff <- post_beta0_diff %>% tibble::rownames_to_column(var="WhichVar") %>%
  mutate(WhichVar = factor(WhichVar, levels=c("Pstd:I(Hstd^2)", "Hstd:I(Pstd^2)",
                                              "I(Hstd^2)","I(Pstd^2)", "Pstd:Hstd", "Hstd", "Pstd","(Intercept)" ))) %>%
  mutate(WhichVar = recode(WhichVar, "Pstd:I(Hstd^2)"="P:JC²", "Hstd:I(Pstd^2)"="JC:P²",
         "I(Hstd^2)"="JC²","I(Pstd^2)"="P²", "Pstd:Hstd"="P:JC", "Hstd"="JC",
         "Pstd"="P","(Intercept)"="(Intercept)" ))

post_beta0_err <- post_beta0_err %>% tibble::rownames_to_column(var="WhichVar") %>%
  mutate(WhichVar = factor(WhichVar, levels=c("Pstd:I(Hstd^2)", "Hstd:I(Pstd^2)",
                                              "I(Hstd^2)","I(Pstd^2)", "Pstd:Hstd", "Hstd", "Pstd","(Intercept)" ))) %>%
  mutate(WhichVar = recode(WhichVar, "Pstd:I(Hstd^2)"="P:JC²", "Hstd:I(Pstd^2)"="JC:P²",
         "I(Hstd^2)"="JC²","I(Pstd^2)"="P²", "Pstd:Hstd"="P:JC", "Hstd"="JC",
         "Pstd"="P","(Intercept)"="(Intercept)" ))

#lidar data
p_diff_lidar <- ggplot(post_beta1_diff) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = 2) +
  aes(x=WhichVar, y =mean, ymin = `0.025quant`, ymax = `0.975quant`) +
  xlab("covariate")+ ylab("Effect size")+
  ggtitle("Posteriors for \u0394height (LiDAR data)")+
  theme_classic() +
  coord_flip()
#p_diff_lidar

p_err_lidar <- ggplot(post_beta1_err) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = 2) +
  aes(x=WhichVar, y =mean, ymin = `0.025quant`, ymax = `0.975quant`) +
  xlab("covariate")+ ylab("Effect size")+
  ggtitle("Posteriors for CV(\u0394height) (LiDAR data)")+
  theme_classic() +
  coord_flip()
#p_err_lidar

#IBM model
p_diff_ibm <- ggplot(post_beta0_diff) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = 2) +
  aes(x=WhichVar, y =mean, ymin = `0.025quant`, ymax = `0.975quant`) +
  xlab("covariate")+ ylab("Effect size")+
  ggtitle("Posteriors for volume (IBM data)")+
  theme_classic() +
  coord_flip()
#p_diff_ibm

p_err_ibm <- ggplot(post_beta0_err) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = 2) +
  aes(x=WhichVar, y =mean, ymin = `0.025quant`, ymax = `0.975quant`) +
  xlab("covariate")+ ylab("Effect size")+
  ggtitle("Posteriors for CV(volume) (IBM data)")+
  theme_classic() +
  coord_flip()
#p_err_ibm

grid.arrange(p_diff_ibm, p_err_ibm, p_diff_lidar, p_err_lidar)
```

#Visualisation of the interpolations

3D graphs for both the model and lidar data  

Difference (height/volume): lijkt me bij combinate hoge P en H en kleine P en H enkel verschillend van elkaar. Bij lidar data gaat bij hoge P en H gaat difference ferm naar beneden, bij IBM data niet. Bij lage P en H gaat bij ibm data de difference naar omhoog, bij lidar data weer naar beneden. Bij de niet extreme waarden lijkt het wel goed overeen te komen (zou eens apart kunnen geplot worden).  

Error: zadelprofiel bij ibm data, bij lidar data niet aanwezig (grote en kleine waarden van H)...

```{r}
#3Dplot
#Lidar data
Interlidar_diff_3D <- plot_ly(data = interlidar_data, x = ~H, y = ~Pstd, z=~Int_diff_mean, color=~Int_diff_mean)  %>% add_markers() %>%
  layout(title = "Difference in dune height (LiDAR data)",
  scene = list(
    xaxis = list(title = "JC"),
    yaxis = list(title = "P", tickvals=c(0,0.2,0.4,0.6,0.8,1)),
    zaxis = list(title = "\u0394height (m)", titleangle=90))) %>%
  colorbar(title = "Prediction \u0394height")


Interlidar_err_3D <- plot_ly(data = interlidar_data, x = ~H, y = ~Pstd, z=~exp(Int_err_mean), color=~exp(Int_err_mean))  %>% add_markers() %>%
  layout(title = "Variability in dune height (LiDAR data)",
  scene = list(
    xaxis = list(title = "JC"),
    yaxis = list(title = "P", tickvals=c(0,0.2,0.4,0.6,0.8,1)),
    zaxis = list(title = "CV (m)"))) %>%
  colorbar(title = "Prediction CV")



#IBM model
Interibm_diff_3D <- plot_ly(data = interibm_data, x = ~H, y = ~Pstd, z=~Int_diff_mean, color=~Int_diff_mean)  %>% add_markers() %>%
  layout(title = "Difference in dune volume (IBM data)",
  scene = list(
    xaxis = list(title = "JC"),
    yaxis = list(title = "P", tickvals=c(0,0.2,0.4,0.6,0.8,1)),
    zaxis = list(title = "volume (m³)"))) %>%
  colorbar(title = "Prediction volume")


Interibm_err_3D <- plot_ly(data = interibm_data, x = ~H, y = ~Pstd, z=~exp(Int_err_mean), color=~exp(Int_err_mean))  %>% add_markers() %>%
  layout(title = "Variability in dune volume (IBM data)",
  scene = list(
    xaxis = list(title = "JC"),
    yaxis = list(title = "P", tickvals=c(0,0.2,0.4,0.6,0.8,1)),
    zaxis = list(title = "CV (m³)"))) %>%
  colorbar(title = "Prediction CV")



Interlidar_diff_3D
Interibm_diff_3D


Interlidar_err_3D
Interibm_err_3D
```

#Plotting uncertainties

```{r}
#3Dplot
#Lidar data
Interlidar_diff_3D_uncertainty <- plot_ly(data = interlidar_data, x = ~H, y = ~Pstd, z=~Int_diff_mean, color=~(Int_diff_0.975quant - Int_diff_0.025quant))  %>% add_markers() %>%
  layout(title = "Uncertainty Difference (lidar data)",
  scene = list(
    xaxis = list(title = "JC"),
    yaxis = list(title = "P", tickvals=c(0,0.2,0.4,0.6,0.8,1)),
    zaxis = list(title = "Height (m)")))

Interlidar_err_3D_uncertainty <- plot_ly(data = interlidar_data, x = ~H, y = ~Pstd, z=~exp(Int_err_mean), color=~exp(Int_err_0.975quant-Int_err_0.025quant))  %>% add_markers() %>%
  layout(title = "Uncertainty Variability (lidar data)",
  scene = list(
    xaxis = list(title = "JC"),
    yaxis = list(title = "P"),
    zaxis = list(title = "Error (m)")))

Interibm_diff_3D_uncertainty <- plot_ly(data = interibm_data, x = ~H, y = ~Pstd, z=~Int_diff_mean, color=~(Int_diff_0.975quant - Int_diff_0.025quant))  %>% add_markers() %>%
  layout(title = "Uncertainty volume (ibm data)",
  scene = list(
    xaxis = list(title = "JC"),
    yaxis = list(title = "P"),
    zaxis = list(title = "Height (m)")))

Interibm_err_3D_uncertainty <- plot_ly(data = interibm_data, x = ~H, y = ~Pstd, z=~exp(Int_err_mean), color=~exp(Int_err_0.975quant-Int_err_0.025quant))  %>% add_markers() %>%
  layout(title = "Uncertainty variability (ibm data)",
  scene = list(
    xaxis = list(title = "JC"),
    yaxis = list(title = "P"),
    zaxis = list(title = "Error (m)")))


Interlidar_diff_3D_uncertainty
Interibm_diff_3D_uncertainty

Interlidar_err_3D_uncertainty
Interibm_err_3D_uncertainty
```

